Left-Linear Rewriting in Adhesive Categories

### 2024-07-02 concur/final.tex
coffee new

  \caption{A sequence of rewriting steps in .}
  \label{fi:coffee-rewriting}
\end{figure}

Just to get some basic intuition about left-linear rewriting systems, their capabilities and the problems which can arise, assume we want to model a situation in which a coffee comes with a bag of sugar, a portion of milk and bag of cocoa. The situation is represented by the graph $G_0$ in Fig.~\ref{fi:coffee-rewriting}: the loop $\mathit{cf}$ is the coffee and the three edges $m$, $s$ and $c$  represent the milk, sugar and cocoa, respectively. Adding one of these elements, call it $x$, in the coffee is modeled as merging the corresponding node with the coffee node so that a self-loop $x$ is created.

The available rules are in Fig.~\ref{fi:coffee-rules}. Rules $\rho_m$ models the addition of milk and sugar, while $\rho_c$ adds cocoa and sugar to the coffee. Once the sugar has been added, the coffee can be stirred and then it is ready to drink. This is represented by rule $\rho_r$ which check for the presence of sugar in the coffee, and stirs it.

A possible rewriting sequence is depicted in Fig.~\ref{fi:coffee-rewriting}, applying $\rho_m$, $\rho_c$ and $\rho_r$ in sequence, thus producing a coffee with milk, sugar and cocoa, ready for drinking. A first interesting observation is that after applying rule $\rho_m$ to $G_0$, thus producing graph $G_1$, we can still apply rule $\rho_c$, with a non-injective match. Intuitively, the two rules do not interfere. Rather, $\rho_c$ applied after $\rho_m$ will have no effect on the sugar edge (intuitively, because sugar has been already added). Indeed, the rules could be also be applied in reverse order. Should they be considered independent? Moreover, since rule $\rho_r$ just require the presence of sugar, it could be anticipated, and applied after $r_m$. This might lead to think that $\rho_c$ and $\rho_m$ are independent. However, this is not completely clear, since in absence of $\rho_m$, the application of $\rho_c$ would be essential to enable $\rho_r$. ``There is great chaos under heaven, [hence] the situation is excellent!''

--- new version (according to latest mail, rules not yet up to date) ---

>>>>

  \caption{A sequence of rewriting steps.}
  \label{fi:coffee-rewriting}
\end{figure}

% Just to get some basic intuition about left-linear rewriting systems, their capabilities and the problems which can arise, assume we want to model a situation in which a coffee comes with a bag of sugar, a portion of milk and bag of cocoa. The situation is represented by the graph $G_0$ in Fig.~\ref{fi:coffee-rewriting}: the loop $\mathit{cf}$ is the coffee and the three edges $m$, $s$ and $c$  represent the milk, sugar and cocoa, respectively. Adding one of these elements, call it $x$, in the coffee is modeled as merging the corresponding node with the coffee node so that a self-loop $x$ is created.

% The available rules are in Fig.~\ref{fi:coffee-rules}. Rules $\rho_m$ models the addition of milk and sugar, while $\rho_c$ adds cocoa and sugar to the coffee. Once the sugar has been added, the coffee can be stirred and then it is ready to drink. This is represented by rule $\rho_r$ which check for the presence of sugar in the coffee, and stirs it.

% A possible rewriting sequence is depicted in Fig.~\ref{fi:coffee-rewriting}, applying $\rho_m$, $\rho_c$ and $\rho_r$ in sequence, thus producing a coffee with milk, sugar and cocoa, ready for drinking. A first interesting observation is that after applying rule $\rho_m$ to $G_0$, thus producing graph $G_1$, we can still apply rule $\rho_c$, with a non-injective match. Intuitively, the two rules do not interfere. Rather, $\rho_c$ applied after $\rho_m$ will have no effect on the sugar edge (intuitively, because sugar has been already added). Indeed, the rules could be also be applied in reverse order. Should they be considered independent? Moreover, since rule $\rho_r$ just require the presence of sugar, it could be anticipated, and applied after $r_m$. This might lead to think that $\rho_c$ and $\rho_m$ are independent. However, this is not completely clear, since in absence of $\rho_m$, the application of $\rho_c$ would be essential to enable $\rho_r$. ``There is great chaos under heaven, [hence] the situation is excellent!''
